---
title: "Assignment"
author: "Deekshu Freelancer"
date: "2023-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,message=FALSE,warning=FALSE}
library(tidyverse)
library(lubridate)
```

### Reading the data
```{r}
df = read.csv("2019-Oct.csv")
df
```

### Frequency tables of categorical variables

```{r}
table(df$event_type)
```

```{r}
table(df$category_code)
```


```{r}
table(df$brand)
```


### Visualisation of frequency tables

```{r}
count_data <- df %>%
  group_by(event_type) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  mutate(event_type = factor(event_type, levels = event_type))

ggplot(data = count_data, aes(x = event_type, y = count)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "Event Type Counts",
       x = "Event Type",
       y = "Count") +
  theme_minimal() + geom_text(aes(label=count), vjust=1.6, color="black",
            position = position_dodge(0.9), size=3.5)
```


```{r}
count_data <- df %>%
  group_by(category_code) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  mutate(category_code = factor(category_code, levels = category_code))

ggplot(data = count_data, aes(x = category_code, y = count)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "category code Counts",
       x = "category code",
       y = "Count") +
  theme_minimal() + geom_text(aes(label=count),color="black")+ coord_flip()
```


```{r}
count_data <- df %>%
  group_by(brand) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  mutate(brand = factor(brand, levels = brand))

ggplot(data = count_data, aes(x = brand, y = count,)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "brand Type Counts",
       x = "brand",
       y = "Count") +
  theme_minimal() + coord_flip()
```

```{r}
count_data <- df %>%
  group_by(brand) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  mutate(brand = factor(brand, levels = brand)) %>% head(10)

ggplot(data = count_data, aes(x = brand, y = count)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "brand Type Counts ( top 10 )",
       x = "brand",
       y = "Count") +
  theme_minimal() + coord_flip()
```

### Price Distrivution 
```{r}
ggplot(df, aes(x = price)) +
  geom_histogram(binwidth = 10, fill = "blue", color = "black") +
  labs(title = "Price Distribution", x = "Price", y = "Frequency") +
  theme_minimal()
```
The distribution of price is rightly skewed

### Summary Statistics

```{r}
summary(df)
```

### Data structure
```{r}
str(df)
```


### inappropriate values

The variables category_code and brand have inappropriate values like "" which are empty. I will be segregating brand to 0 and 1 wth "" as 0 and rest as 1 . Similary for category_code I'm labelling as Unknown , appliances,accessories,furniture and stationery. it also has sport.diving label which I'll be add into unknown as it has only 2 values from the total data so create a dummy at the time of building a model wont make sense.

```{r}

df$category_code[df$category_code == ""] <- "unknown"

merge_categories <- function(category) {
  if (grepl("^furniture", category))
    return("furniture")

  if (grepl("^appliances", category))
    return("appliances")
  
  if (grepl("^accessories", category))
    return("accessories")
  
  if (grepl("^sport", category))
    return("unknown")
  
  return(category)
}

df$category_code <- sapply(df$category_code, merge_categories)

df$brand = ifelse(df$brand == "",0,1)
```


### Oultier detection and removal
```{r}
Q1 <- quantile(df$price, 0.25, na.rm = TRUE)
Q3 <- quantile(df$price, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR

outliers <- df[df$price < lower_bound | df$price > upper_bound, ]

df_no_outliers <- df[df$price >= lower_bound & df$price <= upper_bound, ]

```

### Data structure after outlier removal
```{r}
str(df_no_outliers)
```
### Visualisation after outlier removal
### Distribution of Price
```{r}
ggplot(df_no_outliers, aes(x = price)) +
  geom_histogram(fill = "blue", color = "black") +
  labs(title = "Price Distribution", x = "Price", y = "Frequency") +
  theme_minimal()
```
### Event type count visualisation after outlier removal
```{r}
count_data <- df_no_outliers %>%
  group_by(event_type) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  mutate(event_type = factor(event_type, levels = event_type))

ggplot(data = count_data, aes(x = event_type, y = count)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "Event Type Counts",
       x = "Event Type",
       y = "Count") +
  theme_minimal() + geom_text(aes(label=count), vjust=1.6, color="black",
            position = position_dodge(0.9), size=3.5)
```
### Brand type count visualisation after outlier removal

#### 1 = Top known brands ,  0 = local / unknown brands
```{r}
count_data <- df_no_outliers %>%
  group_by(brand) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  mutate(brand = factor(brand, levels = brand))

ggplot(data = count_data, aes(x = brand, y = count)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "brand Type Counts",
       x = "brand Type",
       y = "Count") +
  theme_minimal() + geom_text(aes(label=count), vjust=1.6, color="black",
            position = position_dodge(0.9), size=3.5)
```

### Category code type count visualisation after outlier removal
```{r}
count_data <- df_no_outliers %>%
  group_by(category_code) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  mutate(category_code = factor(category_code, levels = category_code))

ggplot(data = count_data, aes(x = category_code, y = count)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "category_code Type Counts",
       x = "category_code Type",
       y = "Count") +
  theme_minimal() + geom_text(aes(label=count), vjust=1.6, color="black",
            position = position_dodge(0.9), size=3.5)
```

### Creation of new variables for EDA

```{r}
df_no_outliers$event_time = ymd_hms(df_no_outliers$event_time)
df_no_outliers$day_of_week <- weekdays(df_no_outliers$event_time)
df_no_outliers$hour <- hour(df_no_outliers$event_time)
df_no_outliers$time_of_day <- cut(
  df_no_outliers$hour,
  breaks = c(0, 12, 18, 24),
  labels = c("morning", "afternoon", "evening"),
  include.lowest = TRUE,
  right = FALSE
)
```



### Dummy variables 

```{r}
library(fastDummies)
df = dummy_cols(df_no_outliers, select_columns = c("event_type","category_code"),remove_selected_columns = TRUE)
```

So the dependent variable is 'event_type_cart'


### Exploratory data analysis

### Dealing with event types with time of the day
```{r}
df %>% group_by(event_type_cart,time_of_day) %>% summarise(n=n())
df %>% group_by(event_type_purchase,time_of_day) %>% summarise(n=n())
df %>% group_by(event_type_remove_from_cart,time_of_day) %>% summarise(n=n())
df %>% group_by(event_type_view,time_of_day) %>% summarise(n=n())
```

### Count of event_type_cart with the brand

```{r}
df %>% group_by(brand,event_type_cart) %>% summarise(n=n()) %>% arrange(desc(n))
```

### Finding the top user ids, sessions,product id and category id .
```{r}
df %>% group_by(product_id) %>% summarise(n=n()) %>% arrange(desc(n))
df %>% group_by(category_id) %>% summarise(n=n()) %>% arrange(desc(n))
df %>% group_by(user_id) %>% summarise(n=n()) %>% arrange(desc(n))
df %>% group_by(user_session) %>% summarise(n=n()) %>% arrange(desc(n))
```


```{r}
df %>% filter(event_type_cart == 1) %>% group_by(product_id,event_type_cart) %>% summarise(n=n()) %>% arrange(desc(n))


df %>% filter(event_type_cart == 1) %>% group_by(category_id,event_type_cart) %>% summarise(n=n()) %>% arrange(desc(n))

df %>% filter(event_type_cart == 1) %>% group_by(user_id,event_type_cart) %>% summarise(n=n()) %>% arrange(desc(n))

df %>% filter(event_type_cart == 1) %>% group_by(user_session,event_type_cart) %>% summarise(n=n()) %>% arrange(desc(n))
```

### Finding the average price for dependent variable ( Yes / no)
```{r}
df %>% group_by(event_type_cart) %>% summarise(n = mean(price))
```

### Count of event type cart : dependent variable
```{r}
table(df$event_type_cart)
```


I'll be using ROSE

```{r}
df1 = df
df1 = df1 %>% select(c(category_code_accessories,category_code_apparel.glove,category_code_furniture,category_code_appliances,category_code_stationery.cartrige,category_code_unknown,brand,price,event_type_cart))
```

```{r}
df1 = df1 %>% mutate_if(is.integer,as.factor)
df1$brand = as.factor(df1$brand)
```


```{r}
library(caret)
library(ROSE)


set.seed(123)
index <- createDataPartition(df1$event_type_cart, p = 0.8, list = FALSE)
train_data <- df1[index, ]
test_data <- df1[-index, ]

data.rose <- ROSE(event_type_cart ~ ., data = train_data, seed = 1)$data
table(data.rose$event_type_cart)

ctrl <- trainControl(method = "cv", number = 5, classProbs = FALSE)
model_both <- train(event_type_cart ~ ., data = data.rose, method = "glm", trControl = ctrl)


predictions_both <- predict(model_both, newdata = test_data)

conf_matrix_both <- confusionMatrix(predictions_both, test_data$event_type_cart)
print(conf_matrix_both)
```


### Random Forest
```{r}
library(rpart)
model_dt <- train(event_type_cart ~ ., data = data.rose, method = "rpart", trControl = ctrl)

predictions_dt <- predict(model_dt, newdata = test_data)

conf_matrix_dt <- confusionMatrix(predictions_dt, test_data$event_type_cart)
print(conf_matrix_dt)
```

